---
title: "manipulate"
author: "Anya Mueller"
date: "21/06/2021"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: "show"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

website: https://noaa-iea.github.io/r3-train/manipulate.html
slides: https://docs.google.com/presentation/d/1lilrIzlD5Y2vPEax4fao8GqS-TsxSzNMfyGPgKr7TlM/edit

Adam text editor = made by github folks, good for looking at merge conflicts

icons = rpackage for putting icons into rmarkdown

try to participate in the github community! (via forks and pull requests)

data science is mostly janitor work, its usually very easy to make a model, but hard to get data into the right form

tidy data: 
each variable is in its own column, each observation is in its own row

be able to go between wide and long format

pipes increase readability - very good!

R for data science = the bible for data science

readxl - for reading excel spreadsheets

NOTE: there is a visual editor for Rmarkdown files

## Read online table <- heading 

### Download table (`*.csv`) <- subheading

```{r}
# set variables
csv_url  <- "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_AC.csv"
dir_data <- "data"

# derived variables
csv <- file.path(dir_data, basename(csv_url))

# create directory
dir.create(dir_data)

# download file
download.file(csv_url, csv)
```

### Read table `read.csv()`
```{r}
# attempt to read csv
d <- read.csv(csv)

# show the data frame
d #the second line is units, we don't want that

#convert to tibble
tibble::tibble(d)

# read csv by skipping first two lines, so no header
d <- read.csv(csv, skip = 2, header = FALSE)
d

# update data frame to original column names
names(d) <- names(read.csv(csv))
d

#converto to tibble
tibble::tibble(d)
```

### Show table `DT::datatable()`

```{r}
# show table
DT::datatable(d) # make interactive data table
```

### Manipulate with `dplyr`

```{r}
library(DT)
library(dplyr)

d <- d %>% 
  # tibble
  tibble() %>% 
  # mutate time
  mutate(
    time = as.Date(substr(time, 1, 10))) %>% 
  # select columns
  select(
    time, 
    starts_with("total_fisheries_revenue")) %>% 
  # filter rows
  filter(
    time >= as.Date("1981-01-01"))

datatable(d)
```

### Tidy with `tidyr`

```{r}
library(tidyr)

d <- d %>% 
  pivot_longer(-time) #observations to be unique by time

datatable(d)
```

### Summarize with `dplyr`

```{r}
library(stringr) 

d <- d %>% 
  mutate(
    region = str_replace(name, "total_fisheries_revenue_", "")) %>% 
  select(time, region, value)
datatable(d)

d_sum <- d %>% 
  group_by(region) %>% 
  summarize(
    avg_revenue = mean(value))
datatable(d_sum) %>% 
  formatCurrency("avg_revenue") #format into currency
```

### Apply functions with `purrr` on a `nest`'ed `tibble`

Asking the question : What’s the trend over time for fishing revenue by region?

```{r}
library(purrr) #we are going to nest a bunch of content

n <- d %>% 
  group_by(region) %>% 
  nest(
    data = c(time, value)) #make a new list column called data
n #can see that there is a tibble in the data columns

n <- n %>% #make a linear model per region of data
  mutate(
    lm    = map(data, function(d){ #map takes the input of data, puts it into a fucntion, returns a list
      #browser() <- lets us look inside the function
      lm(value ~ time, d) } ),
    trend = map_dbl(lm, function(m){
      coef(summary(m))["time","Estimate"] })) # make new column with linear model, extract trend from linear model
n
n %>% 
  select(region, trend) %>% 
  datatable()
```

