---
title: "visualize"
author: "Anya Mueller"
date: "28/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) #knitr is for rendering rmark down
```

website: https://noaa-iea.github.io/r3-train/visualize.html

its going to be a brief overview of graphing

we are going to look at ggplot2 (static graph), plotly (interactive), dygraphs (interactive)

ggplot2 - look at cheat sheets! (help/cheat sheets)

```{r Read Data}
# install packages
#install.packages("here")
#install.packages("readr")
#install.packages("DT")

# libraries
library(here)
library(readr)
library(DT)

# variables
url_ac <- "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_AC.csv"
csv_ac <- here("data/cciea_AC.csv")

# read data
d_ac        <- read_csv(url_ac, col_names = F, skip = 2) #skip firt two lines bc there was header and units of header
names(d_ac) <- names(read_csv(url_ac)) #take names from header

# show data
datatable(d_ac)
```

## Plot statically with ggplot2
^this is a second level header

### simple line plot (geom_line)
^this is a thrid level header
```{r}
library(dplyr)
library(ggplot2)

# subset data
d_coast <- d_ac %>% 
  # select columns
  select(time, total_fisheries_revenue_coastwide) %>% 
  # filter rows
  filter(!is.na(total_fisheries_revenue_coastwide))
datatable(d_coast)

# ggplot object
p_coast <- d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = time, y = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_line()

# show plot
p_coast
```

### trend line (geom_smooth)
```{r}
p_coast + 
  geom_smooth(method = "lm")
#other methods
p_coast + 
  geom_smooth(method = "glm")

p_coast + 
  geom_smooth(method = "gam")
```

### distribution of values (geom_histogram)
```{r}
d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_histogram(binwidth = 100) #auto is 30
```

### series line plot (aes colour = region)
need to go back to original data and tidy it up (make long format)
```{r}
library(stringr)
library(tidyr)

d_rgn <- d_ac %>% 
  # select columns
  select(
    time, 
    starts_with("total_fisheries_revenue")) %>% 
  # exclude column
  select(-total_fisheries_revenue_coastwide) %>% 
  # pivot longer
  pivot_longer(-time) %>% #pivot everything but time
  # mutate region by stripping other
  mutate(
    region = name %>% 
      str_replace("total_fisheries_revenue_", "") %>% 
      str_to_upper()) %>% 
  # filter for not NA
  filter(!is.na(value)) %>% 
  # select columns
  select(time, region, value)
  
# create plot object
p_rgn <- ggplot(
  d_rgn,
  # aesthetics
  aes(
    x     = time, 
    y     = value,
    group = region,
    color = region)) +
  # geometry
  geom_line()

# show plot
p_rgn
```

### update labels (labs)
```{r}
p_rgn <- p_rgn +
  labs(
    title  = "Fisheries Revenue",
    x      = "Year",
    y      = "Millions $ (year 2015)",
    color  = "Region")
p_rgn
```

### multiple plots with facet_wrap
usually would do this with another variable (ex species)
```{r}
p_rgn +
  facet_wrap(vars(region))
```

### bar plot (geom_col)
```{r}
library(glue)
library(lubridate)

yr_max <- year(max(d_rgn$time))

d_rgn %>% 
  # filter by most recent time
  filter(year(time) == yr_max) %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_col() +
  # add labels
  labs(
    title  = glue("Fisheries Revenue for {yr_max}"),
    x      = "Region",
    y      = "Millions $ (year 2015)",
    fill   = "Region")
```

### variation of series with geom_boxplot
```{r}
d_rgn %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_boxplot() +
  # add labels
  labs(
    title  = "Fisheries Revenue Variability",
    x      = "Region",
    y      = "Millions $ (year 2015)") +
  # drop legend since redundant with x axis
  theme(
    legend.position = "none") #don't give a legend
```

### variation of series with geom_violin
these show the distribution better
```{r}
p_rgn_violin <- d_rgn %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_violin() +
  # add labels
  labs(
    title  = "Fisheries Revenue Variability",
    x      = "Region",
    y      = "Millions $ (year 2015)") +
  # drop legend since redundant with x axis
  theme(
    legend.position = "none")
p_rgn_violin
```
### change theme 
```{r}
p_rgn_violin +
  theme_classic()
```
you can make a custom theme! (create a function -> theme_anya <- function())

## Plot interactively with plotly or dygraphs
### make ggplot interactive with plotly::ggplotly
this will make an interactive plot for when it is knit into an html object
-VERY COOL! (can check out examples here http://www.htmlwidgets.org/)
```{r}
#install.packages("plotly")
plotly::ggplotly(p_rgn)
```

### create interactive time series with dygraphs::dygraphs
this likes wide data
simplified view for time series
```{r}
#install.packages("dygraphs")
library(dygraphs)
#make data wide
d_rgn_wide <- d_rgn %>% 
  mutate(
    Year = year(time)) %>% 
  select(Year, region, value) %>% 
  pivot_wider(
    names_from  = region,
    values_from = value)
datatable(d_rgn_wide)

d_rgn_wide %>% 
  dygraph() %>% 
  dyRangeSelector()
```

